---
title: "SAMC Comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SAMC Comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Setup

```{r install, eval=FALSE}
devtools::install_github("LandSciTech/LSTDConnect")
```

We load the package along with `raster` and `samc`. We also load the `microbenchmark` package to benchmark code performance. We will use ggplot for some visualizations.

```{r load_libs}
library(LSTDConnect)
library(raster)
library(samc)
library(microbenchmark)
library(ggplot2)
```

Then, access the buolt-in dataset, `ghm`. This raster represent habitat resistance for a small part of Canada (100 cells squares, at a resolution of 1 km).

```{r get_data}
res <- LSTDConnect::ghm
plot(res)
```

### Running SAMC in paralell

`LSTDconnect` is optimized to obtain the transient distributions from a SAMC calculation. We will ise the `ghm` layer as resistance and will use the opposite layer for occupancy `occ` such as `occ = 1-res`.

```{r get_occ}
occ <- 101 - res
plot(occ)
```

Mortality is built from resistance so that when resistance is larger than 80, mortality is high whereas under 80, it is low.

```{r get_mort}
high_mort <- 0.25
low_mort <- 0.01
cutoff <- 80

mort <- res
mort[mort >= cutoff] <- high_mort
mort[mort < cutoff] <- low_mort
```

We first run `LSTDConnect::samc` to prepare the data for the computation.

```{r step_1}
samc_cache <- LSTDConnect::samc(resistance = res, absorption = mort, 
                                directions = 8)
```

We then use `LSTDConnect::distribution`, analogous to `samc::distribution`, to get the transient distribution of individuals for a given time step. We get a 2 elements list, which can be 

```{r step_2}
dists <- LSTDConnect::distribution(samc = samc_cache, occ = occ, time = 1000)
plot(dists$occ)
```

### Comparison with the original samc package

We can compare the results with those given by the `samc` package.

```{r samc_samc}
tr_list <- list(fun = function(x) 1 / mean(x),
                dir = 8, 
                sym = TRUE)

samc_cache_comp <- samc::samc(data = res, absorption = mort, tr_args = tr_list)
dists_comp <- samc::distribution(samc = samc_cache_comp, occ = occ, time = 1000)
dists_mapped <- samc::map(samc_cache_comp, dists_comp)
plot(dists_mapped)
```

Let's look at the differences. We can see the error is extremely small

```{r comp}
diff <- dists_mapped - dists$occ
plot(diff)
```

```{r comp_hist}
hist(values(diff), main = "Error")
```

Let's now benchmark and compare the speed of these two packages by running each process 100 times

```{r}
t <- 1000

samc_obj <- samc::samc(data = res,
                       absorption = mort,
                       tr_args = tr_list)  
samc_obj_custom <- LSTDConnect::samc(resistance = res, 
                                     absorption = mort, 
                                     directions = 8)

mbm <- microbenchmark(
  "samc" = {
    short_disp <- samc::distribution(samc = samc_obj,
                                     occ = mort,
                                     time = t)
    gc()
  },
  "LSTDConnect" = {
    short_disp_custom <- LSTDConnect::distribution(samc = samc_obj_custom,
                                                   occ = occ, 
                                                   time = t)
    gc()
  },
  times = 100,
  unit = "s", 
  control = list(order = "block")) 
mbm
```

```{r benchmark_plot}
boxplot(mbm)
```

